An efficient acyclic contact planner for multiped robots

### 2016-12-13 body/intro.tex
[review] Reviewing introduction.

For this reason previous contributions having demonstrated acyclic contact locomotion on a real robot are too computationally expensive~\citep{Bretl:2006:MPM:1124573.1124585}. Aa a result at the DARPA challenge, the robots fell back to unsafe walking patterns to climb stairs, instead of using the provided handrails to faciliate the motion~\citep{atkensondarpa}.

Obtaining reasonable computation times is critical for a contact planner to be of any practical use. Our work thus aims at breaking the complexity of the acyclic contact planning problem.  To do so we deal sequentially with the two main issues associated to our problem: the null measure of the contact manifold, and the non-determinism of the contact selection problem. First we introduce a low dimensional space, called the \textit{contact-feasible} space, that can be sampled and mapped efficiently to the contact manifold. Then, given a path computed in the \textit{contact-feasible} space, we propose a deterministic algorithm to generate a contact sequence along the path.
This decoupling between the planning of the path and the generation of contacts presents drawbacks and advantages discussed in previous related literature.
%~ whole-body trajectory and the contact location and timings, we decouple path and contact planning, as suggested by \citeauthor{Bouyarmane2009}.
%~ We first plan a path for the geometric root of the robot in a low dimensional space. We formulate this space as a computationally-efficient approximation
%~ of the \textit{equilibrium feasibile} space, that is the space of root configurations for which there exists a joint configuration such that the robot is in static equilibrium.
%~ We then use a deterministic algorithm
%~ This paper presents a geometrical approximation of this condition, and a concrete implementation of such a decoupled planner.
%~ This approximation results from a trade-off between a necessary and a sufficient condition for \textit{equilibrium feasibility}, allowing us to find solutions extremely rapidly, while preserving a high success rate in the demonstrated scenarios.

@@

The issue of planning acyclic contacts was first completely described by Bretl et al. in their seminal paper  \citep{Bretl:2006:MPM:1124573.1124585}. The issue requires the simultaneous handling of two problems, $\mathcal{P}_1$: planning a relevant guide path for the root of the robot in $SE(3)$; and $\mathcal{P}_2$: planning a discrete sequence of acyclic equilibrium configurations along the path. A third nontrivial problem, $\mathcal{P}_3$, 
%~ not addressed in this work, 
then consists in interpolating a complete motion between two postures of the contact sequence.  A key issue is to avoid combinatorial explosion when considering at the same time the possible contacts and the potential paths. This seminal paper proposes a first effective algorithm, able to handle simple situations (such as climbing scenarios), but not applicable to arbitrary environments. Following it, seve\-ral papers have applied this approach in specific situations, limiting the combinatorial by imposing a fixed set of possible contacts \citep{Hauser06usingmotion, stilman2010}.

@@

\textbf{On the other hand, the two problems \Pa and \Pb might be decoupled} to reduce the complexity. The feasibility and interest of the decoupling is shown by \citeauthor{DBLP:conf/iser/EscandeKMG08} who manually set up a rough root guide path (i.e. an ad-hoc solution to $\mathcal{P}_1$). \Pb is addressed as the combinatorial computation of a feasible contact sequence in the neighborhood of the guide. A solution can then be found, %efficiently when considering quadruped locomotion \citep{kalakrishnan2011learning}, 
but at the cost of prohibitive computation times (up to several hours) for constraining scenarios. This approach is suboptimal because the quality of the motion depends on the quality of the guide path. \citeauthor{Bouyarmane2009} precisely focus on automatically computing a guide path with guarantees of \textit{equilibrium feasibility}, by extending key frames of the path into whole-body configurations in static equilibrium. Randomly-sampled configurations are projected into the contact sub-manifold using a inverse-kinematics solver, a computationally-expensive process (about 15 minutes to compute a guide path in the examples presented). Moreover this explicit projection is insufficient to guarantee the feasibility between two key positions in the path. \citeauthor{7140082} also propose a decoupled approach, with a planning phase based on the reachable workspace of the robot limbs, used to judge the ability to make contact with 
a discretized environment. This planning phase does not account for collisions, implying that replanning is required in case of failure. This approach is efficient 
in the demonstrated scenarios. In highly-constraining cases such 
as the car egress scenario that we address here, we believe that including collision constraints in the planning is a requirement.

For completeness, we lastly mention a new kind of approach, recently proposed in the computer graphics field \citep{hamalainen_cpbp_2015}. The authors
use black-box physics simulators to perform Model Predictive Control for the motion of a humanoid character, and manage to obtain dynamically consistent motions
at \gls{interactive} frame rates (we say that a planning method is interactive when the computation time for one contact switch is less than the
time to execute it). While this new approach provides an exciting direction of research, currently the resulting motions
look unnatural, and do not seem applicable to real robots.

As far as robotics applications are concerned, none of the existing planners is able to solve the problem with \gls{interactive} performances.
However, recent contributions to the interpolation between contact poses (problem $\mathcal{P}_3$) have brought promising preliminary solutions \citep{Hauser2014, herzog2015trajectory, Park116, Carpentier2016}. In particular, \citeauthor{Carpentier2016} are able to achieve this with \gls{interactive} performances on a real robot.
Therefore, a planner capable of efficiently solving \Pa and \Pb could outperform all existing planners if coupled with an interpolation method solving $\mathcal{P}_3$.
The main contribution of this paper is exactly this planner.

\subsection{Situation and Contributions}

@@

Our solution belongs the class of decoupled approaches, because we believe that the separation between the generation of the root guide path and the contact sequence is the most promising direction \citep{DBLP:conf/iser/EscandeKMG08} to break the complexity of acyclic contact planning. Compared to previous approaches, our solution has two main novelties:

\noindent \textbf{Regarding $\mathcal{P}_1$}, we propose a fast guide path planning algorithm. The key to its efficiency is that it does not sample
directly the contact manifold, but an approximation of the \textit{contact feasible} space, for which there exists a mapping to the contact manifold.
The space is sampled thanks to the \textit{reachability condition} we introduce, a low-dimensional geometric condition for \textit{contact feasibility}.

\noindent \textbf{Regarding $\mathcal{P}_2$},  we propose a fast method to extend a \textit{contact feasible} path into a sequence of whole-body configurations in static equilibrium. This second step requires the explicit computation of contact configurations. It is guided by dedicated heuristics which fastly synthesize relevant configurations.

This sequential approach is the key to the efficiency of our method, though it
can result in failures:  we show that the \textit{reachability condition} only approximates the \textit{contact feasible} space. However, we demonstrate empirically the validity of the approach: the high success rate, combined with the low computation times, allow us to plan (and re-plan upon failure) multi-contact sequences at \gls{interactive} rates.

As a result, we propose contributions to both problems \Pa and $\mathcal{P}_2$.
\begin{itemize}
\item A low-dimensional, efficient sampling-based approach to plan guide paths.
\item A very efficient and general implementation of an acyclic contact planner, the first one compatible with \gls{interactive} applications.
%~ \item Four heuristics for contact generation. They bias the planner towards configurations that are in robust static equilibrium, or 
%~ more efficient with respect to the task.
\end{itemize}

Finally, to demonstrate the validity of our approach regarding the complete chain, we show that the contact plans we produce can be successfully executed, either in simulation or on the real HRP-2 robot.

>>>>

For this reason previous contributions having demonstrated acyclic contact locomotion on a real robot are too computationally expensive~\citep{Bretl:2006:MPM:1124573.1124585}. Aa a result at the DARPA Robotics Challenge, robots relied on unsafe bipedal walking to climb stairs, instead of using the provided handrails to facilitate the motion~\citep{atkensondarpa}.

Obtaining reasonable computation times is critical for a contact planner to be of any practical use. Our work thus aims at breaking the complexity of the acyclic contact planning problem.  To do so we deal sequentially with the two main issues associated to our problem: the null measure of the contact manifold, and the non-determinism of the contact selection problem. First we introduce a low dimensional space, called the \textit{contact-feasible} space, that can be sampled and mapped efficiently to the contact manifold. Then, given a path computed in the \textit{contact-feasible} space, we propose a deterministic algorithm to generate a contact sequence along the path.
This decoupling \deladp{between the planning of the path and the generation of contacts} presents drawbacks and advantages discussed in previous related literature\addadp{, which we summarize in the following}.
%~ whole-body trajectory and the contact location and timings, we decouple path and contact planning, as suggested by \citeauthor{Bouyarmane2009}.
%~ We first plan a path for the geometric root of the robot in a low dimensional space. We formulate this space as a computationally-efficient approximation
%~ of the \textit{equilibrium feasibile} space, that is the space of root configurations for which there exists a joint configuration such that the robot is in static equilibrium.
%~ We then use a deterministic algorithm
%~ This paper presents a geometrical approximation of this condition, and a concrete implementation of such a decoupled planner.
%~ This approximation results from a trade-off between a necessary and a sufficient condition for \textit{equilibrium feasibility}, allowing us to find solutions extremely rapidly, while preserving a high success rate in the demonstrated scenarios.

@@

\repadp{Acyclic contact planning}{The issue of planning acyclic contacts} was first completely described by Bretl et al. in their seminal paper  \citep{Bretl:2006:MPM:1124573.1124585}. \repadp{This problem}{The issue} requires the simultaneous handling of two \addadp{sub}problems, $\mathcal{P}_1$: planning a relevant guide path for the root of the robot in $SE(3)$; and $\mathcal{P}_2$: planning a discrete sequence of \deladp{acyclic} equilibrium configurations along the path. A third nontrivial problem, $\mathcal{P}_3$, 
%~ not addressed in this work, 
then consists in interpolating a complete motion between two postures of the contact sequence.  A key issue is to avoid combinatorial explosion when considering at the same time the possible contacts and the potential paths. This seminal paper proposes a first effective algorithm, able to handle simple situations (such as climbing scenarios), but not applicable to arbitrary environments. Following it, seve\-ral papers have applied this approach in specific situations, limiting the combinatorial by imposing a fixed set of possible contacts \citep{Hauser06usingmotion, stilman2010}.

@@

\textbf{On the other hand, the two problems \Pa and \Pb might be decoupled} to reduce the complexity. The feasibility and interest of the decoupling is shown by \citeauthor{DBLP:conf/iser/EscandeKMG08} who manually set up a rough root guide path (i.e. an ad-hoc solution to $\mathcal{P}_1$). \Pb is addressed as the combinatorial computation of a feasible contact sequence in the neighborhood of the guide. A solution can then be found, %efficiently when considering quadruped locomotion \citep{kalakrishnan2011learning}, 
but at the cost of prohibitive computation times (up to several hours) for constraining scenarios. This approach is suboptimal because the quality of the motion depends on the quality of the guide path. \citeauthor{Bouyarmane2009} precisely focus on automatically computing a guide path with guarantees of \textit{equilibrium feasibility}, by extending key frames of the path into whole-body configurations in static equilibrium. Randomly-sampled configurations are projected into the contact sub-manifold using a inverse-kinematics solver, a computationally-expensive process (about 15 minutes to compute a guide path in the examples presented). Moreover this explicit projection is insufficient to guarantee the feasibility between two key \repadp{postures}{positions} in the path. \citeauthor{7140082} also propose a decoupled approach, with a planning phase based on the reachable workspace of the robot limbs, used to judge the ability to make contact with 
a discretized environment. This planning phase does not account for collisions, implying that replanning is required in case of failure. This approach is efficient 
in the demonstrated scenarios. In highly-constraining cases such 
as the car egress scenario that we address here, we believe that including collision constraints in the planning is a requirement.

For completeness, we lastly mention a new kind of approach, recently proposed in the computer graphics field \citep{hamalainen_cpbp_2015}. The authors
use black-box physics simulators to perform Model Predictive Control for the motion of a humanoid character, and manage to obtain dynamically consistent motions
at \gls{interactive} frame rates (we say that a \deladp{planning} method is interactive when the computation time for \repadp{planning}{one contact switch} is less than the
time \repadp{for execution}{to execute it}). While this new approach provides an exciting direction of research, currently the resulting motions
look unnatural, and do not seem applicable to real robots.

As far as robotics applications are concerned, none of the existing planners is \deladp{able to solve the problem with} \gls{interactive} \deladp{performances}.
However, recent contributions to the interpolation between contact poses (problem $\mathcal{P}_3$) have brought promising preliminary solutions \citep{Hauser2014, herzog2015trajectory, Park116, Carpentier2016}. In particular, \addadp{the algorithm of} \citeauthor{Carpentier2016} \repadp{is }{are able to achieve this with} \gls{interactive} \deladp{performances on a real robot}.
Therefore, a planner capable of efficiently solving \Pa and \Pb could outperform all existing planners if coupled with an interpolation method solving $\mathcal{P}_3$.
The main contribution of this paper is exactly this planner.

\subsection{\deladp{Situation and} Contributions}

@@

Our solution belongs to the class of decoupled approaches, because we believe \repadp{this to be}{that the separation between the generation of the root guide path and the contact sequence is} the most promising direction~\citep{DBLP:conf/iser/EscandeKMG08} to break the complexity of acyclic contact planning. Compared to previous approaches, our solution has two main novelties:

\noindent \textbf{Regarding $\mathcal{P}_1$}, we propose a fast guide path planning algorithm. The key to its efficiency is that it does not sample
directly the contact manifold, but an approximation of the \textit{contact feasible} space, for which there exists a mapping to the contact manifold. \adnote{It is not clear what the difference is between the contact manifold and the contact feasible space}
The space is sampled thanks to the \textit{reachability condition}\deladp{ we introduce}, a low-dimensional geometric condition for \textit{contact feasibility}.

\noindent \textbf{Regarding $\mathcal{P}_2$},  we propose a fast method to extend a \textit{contact feasible} path into a sequence of whole-body configurations in static equilibrium. This \deladp{second step} requires the explicit computation of contact configurations\repadp{, which we guide}{. It is guided} by dedicated heuristics that quickly synthesize \repadp{feasible}{relevant} configurations.

This sequential approach is the key to the efficiency of our method, though it
can result in failures\deladp{: we show that the \textit{reachability condition} only approximates the \textit{contact feasible} space}. However, we demonstrate empirically the validity of the approach: the high success rate, combined with the low computation times, allow us to plan (and re-plan upon failure) multi-contact sequences at \gls{interactive} rates.

\deladp{As a result, we propose contributions to both problems \Pa and $\mathcal{P}_2$.}
\begin{itemize}
\item \deladp{A low-dimensional, efficient sampling-based approach to plan guide paths.}
\item \deladp{A very efficient and general implementation of an acyclic contact planner, the first one compatible with \gls{interactive} applications.}
%~ \item Four heuristics for contact generation. They bias the planner towards configurations that are in robust static equilibrium, or 
%~ more efficient with respect to the task.
\end{itemize}

Finally, to demonstrate the validity of our approach\deladp{ regarding the complete chain}, we show that the generated contact plans can be successfully executed, either in simulation or on the real HRP-2 robot.
